### 2013-10-20 tex/acknowledge.tex

First, I thank my advisor, Professor Jonathan Goodman, for giving me the opportunity to work on
this problem, and helping me arrive at the following exposition. Next, I wish to thank Professor
Helena Frydman for ﬁrst sparking my interest in Markov chains by giving lucid descriptions of their
many interesting properties and applications. I thank Professors James Demmel and Beresford
Parlett, for answering many questions pertaining to this problem, and Professor Leslie Greengard
for agreeing to review this paper.

Finally, I would like to thank my family, for their support of my interests in research (and
everything else!), and especially my parents, Barbara and Ted Terry, and Benita and Bill DeMeo.
Needless to say, without them this paper would not have been written.